The title

### 2021-08-14 manuscript/manuscript.Rmd
New equalize_trial_counts, io

# Write R package citations
r_refs(here("manuscript", "r-references.bib"), append = FALSE)

# Set global chunk options
opts_chunk$set(
  fig.align = "center",
  fig.width = 10,
  include = FALSE,
  out.width = "100%"
)

# Re-run preprocessing and/or model fitting?
run <- list(preprocessing = FALSE, model_fitting = FALSE)

# Number of CPU cores for parallel model fitting
n_cores <- detectCores()

# File names for easier I/O
deps <- c("logRT", "P1", "N170", "N400")
fnames <- list(
  montage = "results/montage.csv",
  obj_per_cond = "results/obj-per-cond.csv",
  trials = "results/trials.csv",
  ave_grand = "results/ave_grand.csv",
  ave_tfr_flat = "results/ave-tfr_flat.csv",
  ave_tfr_diff = "results/ave-tfr_diff.csv",
  models_lme = "results/models_lme.RDS",
  models_brm = str_c("results/model_brm_", deps, ".RDS") %>% set_names(deps),
  model_brm_prior = "results/model_brm_prior.RDS",
  models_control = "results/models_control.RDS"
)
```

@@

# Generate new BIDS-style subject IDs
seq(length(fname_vhdr)) %>%
  str_pad(width = 2, pad = 0) %>%
  str_c("sub-", .) -> subject_id

# Define ERP components of interest
erp_components <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(0.1, 0.15, 0.4),
  tmax = c(0.15, 0.2, 0.7),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

@@

  # Import the custom Python functions for preprocessing
  pp <- import_from_path("preprocessing", path = "code")

  # Do the actual preprocessing
  tibble(fname_vhdr, fname_log, subject_id) %>%
    pwalk(
      pp$preprocess,
      output_dir = "data/preprocessed",
      eog_names = c("HEOG", "VEOG"),
      eog_anodes = c("F9", "Auge_u"),
      eog_cathodes = c("F10", "Fp1"),
      montage_kind = "easycap-M1",
      ica_components = as.integer(15),
      ica_method = "fastica",
      ica_seed = as.integer(1234),
      hipass = 0.1,
      lowpass = 30,
      event_id = dict(match = as.integer(221), mismatch = as.integer(222)),
      tmin = -0.5,
      tmax = 1.489,
      baseline = tuple(-0.2, 0),
      reject = dict(eeg = 200e-6),
      erp_components = erp_components,
      average_by = c("part", "condition", "subject_id"),
      tfr_baseline = tuple(-0.3, -0.1),
      tfr_equalize = "condition == ['Informed', 'Naive']",
      tfr_resample = py_none(),
      tfr_freqs = tfr_freqs,
      tfr_cycles = tfr_cycles,
      tfr_crop = tuple(-0.2, 0.8)
    )

  # Combine single trial data for all subjects
  dir.create("results")
  str_c("data/preprocessed/", subject_id, "_trials.csv") %>%
    map(read_csv, col_types = cols()) %>%
    bind_rows() %>%
    write_csv(fnames$trials)

  # Get standard EasyCap electrode positions and project them into 2D space
  read_csv("data/preprocessed/sub-01_tfr-ave.csv", col_types = cols()) %>%
    select(-c(part, condition, subject_id, time, freq)) %>%
    colnames() -> ch_names
  import("mne.channels._standard_montage_utils")$MONTAGE_PATH %>%
    str_c("/easycap-M1.txt") %>%
    read_tsv(col_types = cols()) %>%
    rename(electrode = Site) %>%
    inner_join(tibble(electrode = ch_names), ., by = "electrode") %>%
    mutate(
      x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
      y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)),
    ) %>%
    mutate(across(c(x, y), round, 3)) %>%
    write_csv(fnames$montage)

  # Determine number of objects per condition for each participant
  fname_log %>%
    map2(subject_id, pp$read_log) %>%
    bind_rows() %>%
    with(table(subject_id, condition) / 3) %>%
    as.data.frame.matrix() %>%
    rownames_to_column("subject_id") %>%
    write_csv(fnames$obj_per_cond)
}
```

@@

# Average across subjects
id_vars = ["part", "condition", "time"]
ave = ave.groupby(id_vars, as_index=False).mean()
ave.to_csv(r.fnames["ave_grand"], float_format="%.3f", index=False)

@@

# Average across subjects and save
id_vars.remove("subject_id")
dfs = [ave_tfr_flat, ave_tfr_diff]
fnames_tfr = [r.fnames["ave_tfr_flat"], r.fnames["ave_tfr_diff"]]
for df, fname in zip(dfs, fnames_tfr):
    df = df.groupby(id_vars, as_index=False).mean()
    df.to_csv(fname, float_format="%.3f", index=False)
del df, dfs, ave, ave_tfr, ave_tfr_diff, ave_tfr_flat
```

```{r, models_lme}
# Compute average number of objects per condition
(obj_per_cond <- read_csv(fnames$obj_per_cond, col_types = cols()) %>%
  summarise(across(.cols = -subject_id, .fns = mean)) %>%
  as.list())

@@

  # Load models
  models_lme <- readRDS(fnames$models_lme)
} else {

@@

    # Automic model selection via likelihood ratio tests
    # using an alpha level of .20 (see Matuschek et al., 2017, J Mem Lang)
    cl <- makeCluster(n_cores)
    build <- buildmer(
      buildmerControl = buildmerControl(
        formula,
        data,
        direction = c("backward", "backward"),
        cl = cl,
        crit = "LRT",
        elim = function(logp) exp(logp) >= .20,
        calc.anova = TRUE,
        ddf = "Satterthwaite",
        dep = dep,
        REML = FALSE,
      ),
      control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e6))
    )
    stopCluster(cl)

@@

  # Compute models for all dependent variables in parallel
  form <- RT ~ part * condition +
    (part * condition | subject_id) +
    (part * condition | item_id)
  map(
    deps,
    build_lme,
    formula = form,
    data = trials_main,
    n_cores = n_cores
  ) %>%
    set_names(deps) %T>%
    saveRDS(fnames$models_lme) -> models_lme
}
```

@@

# Fit Bayesian mixed model using prior information only
fit_brm_prior <- brm(
  formula = formula_brm,
  data = trials,
  family = gaussian(),
  prior = prior,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 20000,
  warmup = 2000,
  file = fnames$model_brm_prior,
  file_refit = "never"
)

@@

# Compute posterior distributions for dependent variables
deps <- c("logRT", "P1", "N170", "N400")
map(deps, function(dv) {

  # Define model formula for the current dependent variable
  formula_dv <- str_c(
    dv, " ~ part * condition +
      (part * condition | subject_id) +
      (part * condition | item_id)"
  ) %>%
    formula()

  # Sample from the posterior distribution
  fit_brm <- update(
    fit_brm_prior,
    formula = formula_dv,
    newdata = trials,
    sample_prior = "yes",
    file = fnames$models_brm[[dv]],
    file_refit = "never"
  )

@@

```{r, fig1, include=TRUE, fig.height=10, fig.cap="(ref:figure-1-caption)"}
# Load montage and evoked data
montage <- read_csv(fnames$montage, col_types = cols())
evokeds <- read_csv(fnames$ave_grand, col_types = cols(), progress = FALSE)

@@

```{r, fig3, eval=FALSE, include=TRUE}
# Read grand averages of evoked power
ave_tfr_flat <- read_csv(fnames$ave_tfr_flat, col_types = cols(), progress = FALSE)
ave_tfr_diff <- read_csv(fnames$ave_tfr_diff, col_types = cols(), progress = FALSE)

@@

  # Load models
  models_control <- readRDS(fnames$models_control)
} else {

  # Compute models for all dependent variables in parallel
  form <- RT ~ part * condition +
    (part * condition | subject_id) +
    (part * condition | item_id)
  map(
    deps,
    build_lme,
    formula = form,
    data = trials_control,
    n_cores = n_cores
  ) %>%
    set_names(deps) %T>%
    saveRDS(fnames$models_control) -> models_control
}
```

>>>>

# Write R package citations to bibliography
r_refs(here("manuscript", "r-references.bib"), append = FALSE)

# Set global chunk options
opts_chunk$set(fig.width = 10, include = FALSE, out.width = "100%")

# Re-run preprocessing and/or model fitting?
run <- list(preprocessing = TRUE, model_fitting = TRUE)

# File names for easier input/output
deps <- c("logRT", "P1", "N170", "N400")  # dependent variables
io <- list(
  montage = "results/montage.csv",
  obj_per_cond = "results/obj-per-cond.csv",
  trials = "results/trials.csv",
  ave_grand = "results/ave_grand.csv",
  ave_tfr_flat = "results/ave-tfr_flat.csv",
  ave_tfr_diff = "results/ave-tfr_diff.csv",
  models_lme = "results/models_lme.RDS",
  models_brm = str_c("results/model_brm_", deps, ".RDS") %>% set_names(deps),
  model_brm_prior = "results/model_brm_prior.RDS",
  models_control = "results/models_control.RDS"
)
```

@@

# Generate new BIDS-style subject IDs (sub-01, sub-02, ...)
seq_along(fname_vhdr) %>%
  str_pad(width = 2, pad = 0) %>%
  str_c("sub-", .) -> subject_id

# Define ERP components of interest
erp_components <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(0.100, 0.150, 0.400),
  tmax = c(0.150, 0.200, 0.700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

@@

  # Import the custom Python functions for preprocessing
  pp <- import_from_path("preprocessing", path = here("manuscript"))

  # Do the actual preprocessing
  tibble(fname_vhdr, fname_log, subject_id) %>%
    pwalk(
      pp$preprocess,
      output_dir = "data/preprocessed",
      resample_sfreq = 250,
      eog_names = c("HEOG", "VEOG"),
      eog_anodes = c("F9", "Auge_u"),
      eog_cathodes = c("F10", "Fp1"),
      montage_kind = "easycap-M1",
      ica_components = as.integer(15),
      ica_method = "fastica",
      random_seed = as.integer(1234),
      hipass = 0.1,
      lowpass = 30,
      event_id = dict(match = as.integer(221), mismatch = as.integer(222)),
      tmin = -0.5,
      tmax = 1.489,
      baseline = tuple(-0.2, 0),
      reject = dict(eeg = 200e-6),
      erp_components = erp_components,
      average_by = c("part", "condition", "subject_id"),
      tfr_baseline = tuple(-0.3, -0.1),
      tfr_equalize = TRUE,
      condition_col = "condition",
      item_col = "item_id",
      drop_conditions = c("Excl_informed", "Excl_naive", "Excl_known"),
      tfr_freqs = tfr_freqs,
      tfr_cycles = tfr_cycles,
      tfr_crop = tuple(-0.3, 0.7)
    )

  # Combine single trial data for all subjects
  dir.create("results", showWarnings = FALSE)
  str_c("data/preprocessed/", subject_id, "_trials.csv") %>%
    map(read_csv, col_types = cols()) %>%
    bind_rows() %>%
    write_csv(io$trials)

  # Get standard EasyCap electrode positions and project them into 2D space
  read_csv("data/preprocessed/sub-01_tfr-ave.csv", col_types = cols()) %>%
    select(-c(part, condition, subject_id, time, freq)) %>%
    colnames() -> ch_names
  import("mne.channels._standard_montage_utils")$MONTAGE_PATH %>%
    str_c("/easycap-M1.txt") %>%
    read_tsv(col_types = cols()) %>%
    rename(electrode = Site) %>%
    inner_join(tibble(electrode = ch_names), ., by = "electrode") %>%
    mutate(
      x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
      y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)),
    ) %>%
    mutate(across(c(x, y), round, 3)) %>%
    write_csv(io$montage)

  # Determine number of objects per condition for each participant
  fname_log %>%
    map2(subject_id, pp$read_log) %>%
    bind_rows() %>%
    with(table(subject_id, condition) / 3) %>%
    as.data.frame.matrix() %>%
    rownames_to_column("subject_id") %>%
    write_csv(io$obj_per_cond)
}
```

@@

# Average across subjects
id_vars = ["part", "condition", "time"]
ave = ave.groupby(id_vars, as_index=False).mean()
ave.to_csv(r.io["ave_grand"], float_format="%.3f", index=False)

@@

# Average across subjects and save
id_vars.remove("subject_id")
dfs = [ave_tfr_flat, ave_tfr_diff]
fnames_tfr = [r.io["ave_tfr_flat"], r.io["ave_tfr_diff"]]
for df, fname in zip(dfs, fnames_tfr):
    df = df.groupby(id_vars, as_index=False).mean()
    df.to_csv(fname, float_format="%.3f", index=False)
del df, dfs, ave, ave_tfr, ave_tfr_diff, ave_tfr_flat
```

```{r, models_lme}
# Compute average number of objects per condition
(obj_per_cond <- read_csv(io$obj_per_cond, col_types = cols()) %>%
  summarise(across(.cols = -subject_id, .fns = mean)) %>%
  as.list())

@@

  # Load models
  models_lme <- readRDS(io$models_lme)
} else {

@@

    # Automic model selection via likelihood ratio tests
    # using an alpha level of .20 (see Matuschek et al., 2017, J Mem Lang)
    n_cores <- detectCores()
    cl <- makeCluster(n_cores)
    build <- buildmer(
      buildmerControl = buildmerControl(
        formula,
        data,
        direction = c("backward", "backward"),
        cl = cl,
        crit = "LRT",
        elim = function(logp) exp(logp) >= .20,
        calc.anova = TRUE,
        ddf = "Satterthwaite",
        dep = dep,
        REML = FALSE,
      ),
      control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e6))
    )
    stopCluster(cl)

@@

  # Compute models for all dependent variables in parallel
  form <- RT ~ part * condition +
    (part * condition | subject_id) +
    (part * condition | item_id)
  map(
    deps,
    build_lme,
    formula = form,
    data = trials_main,
    n_cores = n_cores
  ) %>%
    set_names(deps) %T>%
    saveRDS(io$models_lme) -> models_lme
}
```

@@

# Fit Bayesian mixed model using prior information only
fit_brm_prior <- brm(
  formula = formula_brm,
  data = trials,
  family = gaussian(),
  prior = prior,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 20000,
  warmup = 2000,
  file = io$model_brm_prior,
  file_refit = "never"
)

@@

# Compute posterior distributions for dependent variables
deps <- c("logRT", "P1", "N170", "N400")
map(deps, function(dep) {

  # Define model formula for the current dependent variable
  formula_dv <- str_c(
    dep, " ~ part * condition +
      (part * condition | subject_id) +
      (part * condition | item_id)"
  ) %>%
    formula()

  # Sample from the posterior distribution
  fit_brm <- update(
    fit_brm_prior,
    formula = formula_dv,
    newdata = trials,
    sample_prior = "yes",
    file = io$models_brm[[dep]],
    file_refit = "never"
  )

@@

```{r, fig1, include=TRUE, fig.height=10, fig.cap="(ref:figure-1-caption)"}
# Load montage and evoked data
montage <- read_csv(io$montage, col_types = cols())
evokeds <- read_csv(io$ave_grand, col_types = cols(), progress = FALSE)

@@

```{r, fig3, eval=FALSE, include=TRUE}
# Read grand averages of evoked power
ave_tfr_flat <- read_csv(io$ave_tfr_flat, col_types = cols(), progress = FALSE)
ave_tfr_diff <- read_csv(io$ave_tfr_diff, col_types = cols(), progress = FALSE)

@@

  # Load models
  models_control <- readRDS(io$models_control)
} else {

  # Compute models for all dependent variables in parallel
  form <- RT ~ part * condition +
    (part * condition | subject_id) +
    (part * condition | item_id)
  map(
    deps,
    build_lme,
    formula = form,
    data = trials_control,
    n_cores = n_cores
  ) %>%
    set_names(deps) %T>%
    saveRDS(io$models_control) -> models_control
}
```
